---
title: "Homework Assignment: Sentiment Analysis of Emma"
author: "Roberta Southwood"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center",
                      fig.height = 3, fig.width = 4)
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

**Title**: Exploring Joyful Language in Jane Austen’s *Emma* using Tidytext

### **Instructions**

Complete the following exercises using the appropriate packages in R. Ensure that your solutions are optimized and use functional programming principles where applicable.

1.  Load the necessary libraries.
2.  Answer each question in separate R code chunks.
3.  Provide detailed explanations for your approach.
4.  Submit the rendered HTML file.

#### **Objective**:

Use the **`tidytext`** package and three different **sentiment lexicons** (`nrc`, `afinn`, `bing`) to explore **positive/joyful words** in *Emma* by Jane Austen. You will tokenize the text, apply sentiment filters, visualize frequent sentiment words using `ggplot2`, and create a word cloud.

### **Tasks**:

1.  **Data Preparation**

    -   Load the `austen_books()` dataset from the **`janeaustenr`** package.
    -   Group by book and detect chapter boundaries using regex.
    -   Create `linenumber` and `chapter` columns.

```{r}

# load libraries 
library(tidyverse)
library(dplyr)
library(textdata)
library(tidytext)
library(stringr)
library(janeaustenr)
library(wordcloud)
```

```{r}

# Load the austen_books() dataset from the janeaustenr package
austen_books() |>
  distinct(book)

# Group by book and detect chapter boundaries using regex.
# Create linenumber and chapter columns. 
Austen <- austen_books() |>
  group_by(book) |>
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
                                                 ignore_case = TRUE)))) |>
  ungroup()

# View results 
head(Austen)
  
```

2.  **Tokenization**

    -   Use `unnest_tokens()` to tokenize text into individual words.

```{r}
# tokenization using unnest_tokens()
tidy_austen <- Austen |>
  unnest_tokens(word, text)

nrow(tidy_austen) #725055
```

```{r}

# remove stop words, to ensure they do not impact analysis later 

tidy_austen <- tidy_austen |>
  anti_join(stop_words)

nrow(tidy_austen) # 217609
```

3.  **Sentiment Analysis**

    -   Filter joy/positive words from **each** of the three sentiment lexicons:

        -   `nrc` (joy)
        -   `afinn` (positive scores ≥ 1)
        -   `bing` (positive)

    -   Join each with *Emma*'s text and:

        -   Count word frequency.
        -   Filter for frequently occurring words (`n > 50`).
        -   Visualize using a **bar chart** (`ggplot2`) and a **word cloud** (`wordcloud`).

```{r}

# Filter to only the book Emma
Emma <- tidy_austen |>
  filter(book == "Emma")

```

**Joy sentiment** `nrc` (joy)

```{r}

# get nrc lexicon 
nrc_lexicon <- get_sentiments("nrc")

# join with Emma 
emma_nrc_joined <- Emma |>
  inner_join(nrc_lexicon)

# filter for nrc joy 
emma_joy_nrc <- emma_nrc_joined |>
  filter(sentiment == "joy")

# count word frequency 
count(emma_joy_nrc) # 3691
```

```{r}
# filter by words that appear more than 50 times and create bar chart 
emma_joy_nrc |>
  count(word, sort =TRUE) |>
  filter(n > 50) |>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(n, word)) +
  geom_col(fill = "purple", color = "black") +
  labs(y = NULL, title = "Top Joy Sentiment Words in Emma (nrc)")

```

```{r}

# filter to words that appear over 50 times 
emma_joy_nrc_over_50 <- emma_joy_nrc |>
  count(word, sort =TRUE) |>
  filter(n > 50) 

# create a word cloud for the words that appear over 50 times 
wordcloud(words = emma_joy_nrc_over_50$word, 
       freq = emma_joy_nrc_over_50$n, 
       scale=c(3, 0.5),
       max.words = 20)


```

**Positive score** `afinn` (positive scores ≥ 1)

```{r}

# get affin lexicon 
afinn_lexicon <- get_sentiments("afinn")

# join affin with emma 
emma_affin <-  Emma |>
  inner_join(afinn_lexicon)

# filter for positive scores >= 1
emma_positive_affin <- emma_affin |>
  filter(value >= 1 )

# count word frequency 
count(emma_positive_affin) # 4518
```

```{r}

# filter by words that appear more than 50 times and create bar chart 
emma_positive_affin |>
  count(word, sort =TRUE) |>
  filter(n > 50) |>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(n, word)) +
  geom_col(fill = "green", color = "black") +
  labs(y = NULL, title = "Top Positive Words in Emma (affin >=1)")

```

```{r}

# filter to words that appear over 50 times 
emma_positive_affin_over_50 <- emma_positive_affin |>
  count(word, sort =TRUE) |>
  filter(n > 50) 

# create a word cloud for the words that appear over 50 times 
wordcloud(words = emma_positive_affin_over_50$word, 
       freq = emma_positive_affin_over_50$n, 
       max.words = 20)
```

**Positive Sentiment** `bing` (positive)

```{r}

# get bing lexicon 
bing_lexicon <- get_sentiments("bing")

# join with emma  
emma_bing <-  Emma |>
  inner_join(bing_lexicon)

# filter for positive sentiment  
emma_positive_bing <- emma_bing |>
  filter(sentiment == "positive")

# count word frequency 
count(emma_positive_bing) # 5150
```

```{r}

# filter by words that appear more than 50 times and create bar chart 
emma_positive_bing |>
  count(word, sort =TRUE) |>
  filter(n > 50) |>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(n, word)) +
  geom_col(fill = "orange", color = "black") +
  labs(y = NULL, title = "Top Positive Words in Emma (bing)")

```

```{r}

# filter to words that appear over 50 times 
emma_positive_bing_over_50 <- emma_positive_bing |>
  count(word, sort =TRUE) |>
  filter(n > 50) 

# create a word cloud for the words that appear over 50 times 
wordcloud(words = emma_positive_bing_over_50$word, 
       freq = emma_positive_bing_over_50$n, 
       scale=c(3, 0.5),
       max.words = 20)


```

4.  **Push to GitHub**

    -   Push your complete rendred qmd (`.html` file) to a **GitHub repository**.

    -   Your file should include:

        -   Data wrangling
        -   Sentiment filtering and joins
        -   Visualization code

    -   You must include **at least 5 meaningful commits** to document your workflow.

------------------------------------------------------------------------

## **Rubric 100 Points**

| Category                                       | Points  |
|------------------------------------------------|---------|
| Step 1: Data wrangling (linenumber, chapter)   | 10      |
| Step 2: Tokenization                           | 10      |
| Step 3a: NRC sentiment analysis + plots        | 10      |
| Step 3b: AFINN sentiment analysis + plots      | 10      |
| Step 3c: BING sentiment analysis + plots       | 10      |
| Code readability, structure, and comments      | 10      |
| Word cloud for each lexicon                    | 20      |
| GitHub submission                              |         |
| Commit history (at least 5 meaningful commits) | 20      |
| **Total**                                      | **100** |

Good Luck!
